Hybrid MILP to efficiently and accuratly \\ solve hard DNN verification instances

### 2024-09-25 ICLR25/experiments.tex

For that, we considered the challenging DNN MNIST $6 \times 500$, and the 20 first images from the MNIST benchmark. We first run the attack from $\alpha,\beta$-Crown for varying $\epsilon$, using a binary search, to set up an upper bound on $\epsilon$ (the initialization is $\epsilon \in [0,1]$).
Then we run binary search with a global time-out of $10 000s$, initialized from $0$ to this upper bound, where each call is either to $\alpha,\beta$-Crown with TO=2000s, or Hybrid MILP.
We report the results (upper bound, best bound verified by $\alpha,\beta$-Crown and by Hybrid MILP) for each image. We report the results in Fig. \ref{fig2}

\begin{figure}[t!]
	\begin{centering}
\includegraphics[scale=0.5]{epsilon.png}.
\caption{$\epsilon$-robustness proved after $10 000s$ for each of the 20 first images of MNIST.}
\label{fig2}
	\end{centering}
\end{figure}

@@

Compared with $\alpha,\beta$-CROWN, Hybrid MILP is much closer to the upper bound, except for 2 cases (images 8 and 18) where $\alpha,\beta$-CROWN is already very close to the upper bound. On average, the $\epsilon$-gap to upper bound is $0.014$ for Hybrid MILP, and $0.042$ for $\alpha,\beta$-CROWN.

>>>>

For that, we considered the challenging DNN MNIST $6 \times 500$, and the 20 first images from the MNIST benchmark. We first run the attack from $\alpha,\beta$-Crown for varying $\epsilon$, using a binary search, to set up an upper bound on $\epsilon$ (the initialization is $\epsilon \in [0,1]$).
Then we run binary search with a global time-out of $10 000s$, initialized from $0$ to this upper bound, where each call is either to $\alpha,\beta$-Crown with TO=2000s, or Hybrid MILP.
We report the results (upper bound, best bound verified by $\alpha,\beta$-Crown and by Hybrid MILP) for each image. We report the results in Fig. \ref{fig2}.

\begin{figure}[t!]
	\begin{centering}
\includegraphics[scale=0.5]{epsilon.png}.
\caption{$\epsilon$-robustness proved after $10 000s$ for $6 \times 500$ on each of the 20 first images of MNIST.}
\label{fig2}
	\end{centering}
\end{figure}

@@

Compared with $\alpha,\beta$-CROWN, Hybrid MILP is much closer to the upper bound, except for 2 cases (images 8 and 18) where $\alpha,\beta$-CROWN is already very close to the upper bound. On average, the $\epsilon$-gap to upper bound is $0.014$ for Hybrid MILP, 3 times smaller than the $0.042$ gap for $\alpha,\beta$-CROWN.

### 2024-09-25 ICLR25/formula.tex
edit section 4

	\begin{align}
		\Utility\_\max\nolimits^z(a) = W_{az} \times (\sol(\hat{a})- \ReLU(\sol(a))).
	\end{align}

@@

Now, consider neurons $a$ two layers before $z$. We use $b$ to denote neurons in the layer $\ell$ between $a$ and $z$. 
If we open $a$ without changing its value $\sol(a)$, then the change $\Delta(\hat{a})$ in the weight of $\hat{a}$ is 
$\Delta(\hat{a})=\ReLU(\sol(a)) - \sol(\hat{a}) \leq 0$ as above. Its impact on $z$ is no more direct with $W_{az}$, but it is through all the $b \in \ell$. 
We let $\Delta(b) = W_{ab}\Delta(\hat{a})$.
Based on Proposition \ref{LP}, we can evaluate the impact 
$\Delta(\hat{b})$ of opening $a$ on each value of $\hat{b}$, by using the upper and lower bound $\UB(b),\LB(b)$:

@@

	It is easy to check that $(\sol'(n))_{n \leq z}$ satisfies the constraints in 
	$\mathcal{M}_{\{a\}}$, as opening $a$ changes the value of $\hat{a}$ from
	$\sol(\hat{a})$ to $\ReLU(\sol(a))$, and the contribution from $a$ to $b,\hat{b},z$ 
	are respected.
	As $\sol'(z)$ is a solution of $\mathcal{M}_{\{a\}}$, it is smaller or equal to the maximal solution: $\sol'(z) \leq$ $\Sol\_\max_{\{a\}}^z(z)$. That is, 
	$\sol(z)-\sol'(z) \geq \sol(z) -$ $\Sol\_\max_{\{a\}}^z(z)$, i.e. 
	$\Improve\_\max^z(a) \leq \Utility\_\max^z(a)$.
	In particular, we have that $\Utility\_\max^z(a) \geq 0$, which was not obvious from the definition.
	\end{proof}

>>>>

	$$\Utility\_\max\nolimits^z(a) = W_{az} \times (\sol(\hat{a})- \ReLU(\sol(a)))$$

@@

Now, consider neuron a $a$ two layers before $z$. We use $b$ to denote neurons in the layer $\ell$ between $a$ and $z$. 
If we open $a$ without changing its value $\sol(a)$, then the change $\Delta(\hat{a})$ in the weight of $\hat{a}$ is 
$\Delta(\hat{a})=\ReLU(\sol(a)) - \sol(\hat{a}) \leq 0$ as above. Its impact on $z$ is no more direct with $W_{az}$, but it is through $\ell$. 
We let $\Delta(b) = W_{ab}\Delta(\hat{a})$ for all $b \in \ell$.
Based on Proposition \ref{LP}, we can evaluate the impact 
$\Delta(\hat{b})$ of opening $a$ on the value of each $\hat{b}$, by using the upper and lower bound $\UB(b),\LB(b)$:

@@

	It is easy to check that $(\sol'(n))_{n \leq z}$ satisfies the constraints in 
	$\mathcal{M}_{\{a\}}$, as opening $a$ changes the value of $\hat{a}$ from
	$\sol(\hat{a})$ to $\ReLU(\sol(a))$, and the contribution from $a$ to $b,\hat{b},z$ 
	are respected.
	As $\sol'(z)$ is a solution of $\mathcal{M}_{\{a\}}$, it is smaller or equal to the maximal solution: $\sol'(z) \leq$ $\Sol\_\max_{\{a\}}^z(z)$. That is, 
	$\sol(z)-\sol'(z) \geq \sol(z) -$ $\Sol\_\max_{\{a\}}^z(z)$, i.e. 
	$ \Utility\_\max^z(a) \geq \Improve\_\max^z(a)$.
	In particular, we have that $\Utility\_\max^z(a) \geq 0$, which was not obvious from the definition.
	\end{proof}
